---
title: "Lyrics Analysis"
output: html_notebook
---

Lets us read the file.

```{r}
library(readr)
lyrics <- read_csv("songdata.csv")
head(lyrics)
```

Lets us examine the dimension of the lyrics dataframe.

```{r}
dim(lyrics)
```

```{r}
library(dplyr)
glimpse(lyrics)
```

Analysis of 55000+ lyrics data
- Number of artists
- Which artist has highest and lowest number of songs 
- Distribution of songs of all artists in the dataset
- Distribution of lyrics length
- Which song lyrics has maximum number of words
- Which song lyrics has minimum number of words
- Distribution of words count in title
- Which songs title has maximum number of words 
- Which songs title has minimum number of words
- WordClouds of titles with minimum and maximum lengths
- Is there a relation between title length and song length?
- Sentiments of the songs (NRC, Bing)
- Which words are most occuring in the lyrics of the songs
- Wordcloud of most popular words in the songs
- Top words used by an artist in his/her songs
- Are there some common Rythmic words that repeats again and again?


Let's start with finding out how many artists are listed in the data. Also, how many songs each artist has.

```{r}
artist<- as.data.frame(table(as.data.frame(lyrics$artist)))
colnames(artist) <- c("artist", "Num_of_songs")
head(artist)
```

Let's see the which artist has most and least number of songs in the dataset.

```{r}
most_songs <- arrange(artist, desc(Num_of_songs))
most_songs
```

```{r fig.width=5, fig.height=3, echo=FALSE}
library(ggplot2)
library(Rmisc)
tilt_theme <- theme(axis.text.x=element_text(angle=45, hjust=1))
p1 <- ggplot(data = head(most_songs,10), aes(artist, Num_of_songs, fill = Num_of_songs)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Num_of_songs), vjust=1.6, color="white", size=3) +
      ggtitle("Artists with most number of songs") +
      tilt_theme
p1
```

```{r}
least_songs <- tail(most_songs, 15)
p2 <- ggplot(data = least_songs, aes(artist, Num_of_songs, fill = Num_of_songs)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=Num_of_songs), vjust=1.6, color="white", size=3) +
      ggtitle("Artists with least number of songs") +
      tilt_theme
p2
```

Let's check the distribution of songs for all artists.

```{r}
p3 <- ggplot(artist, aes(x=Num_of_songs)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="red")
p3
```

Let's analyze the number of words in each song and its distribution.

```{r}
library(stringr)
count_words <- function(vec){
  return (length(unlist((str_extract_all(tolower(vec), '\\w+')))))
}
lyrics$word_count <- sapply(lyrics$text, count_words)
head(lyrics$word_count)
```

```{r}
p4 <- ggplot(lyrics, aes(x=word_count)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="red")
p4
```

Let's analyze the title of the songs, their wordcount and their distribution

```{r}
lyrics$title_word_count <- sapply(lyrics$song, count_words)
head(lyrics$title_word_count)
```

Let's check out the songs that are longest and shortest.

```{r}
longest_song <- arrange(lyrics, desc(word_count))
longest_song <- head(longest_song, 10)
shortest_song <- arrange(lyrics, word_count)
shortest_song <- head(shortest_song, 10)
longest_song
shortest_song
```



```{r}
p5 <- ggplot(data = longest_song, aes(song, word_count, fill = title_word_count)) +
      geom_bar(stat = "identity") +
      geom_text(aes(label=title_word_count), vjust=1.6, color="white", size=3) +
      ggtitle("Longest Songs") +
      tilt_theme
p6 <- ggplot(data = shortest_song, aes(song, word_count, fill = title_word_count)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = title_word_count), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Shortest Songs") +
      tilt_theme
multiplot(p5, p6, cols=2)
```




```{r}
p7 <- ggplot(lyrics, aes(x=title_word_count)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", binwidth = 1, bins = 1)+
 geom_density(alpha=.2, fill="red")
p7
```

WordCloud of popular words from song titles

```{r}
library(wordcloud)
library(SnowballC)
library(RColorBrewer)
library(tm)
texts <- lyrics$song
#texts <- iconv(texts, to = "utf-8")
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords('english'))
corpus <- tm_map(corpus, stemDocument)
corpus <- tm_map(corpus, removeWords, c("and", "this", "there")) 
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
d <- d[-which(d$word %in% c("and","this","that")),]
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```

There are many song titles that are of length 1, 2 and 3. But surprisingly, there are titles of length more than 13 too. Let's check them out.

```{r}
longest_title <- subset(lyrics, lyrics$title_word_count > 13)
longest_title
shortest_title <- subset(lyrics, lyrics$title_word_count == 1)
shortest_title
```

There are 8 songs with title length more than 13 and 8342 songs with single word title. Let's see word cloud of single word titles and longest titles

```{r}
texts <- longest_title$song
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,scale=c(2,0.5),
          max.words=100, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```

```{r}
texts <- shortest_title$song
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,scale=c(2,0.5),
          max.words=100, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```

An interesting questin would be is there relation between length of title and songs? Most probably now, but let's check out.

```{r}
p8 <- ggplot(lyrics, aes(x=factor(title_word_count), y=word_count, fill = factor(title_word_count))) + 
  geom_boxplot() 
p8 
```

```{r}
cor(lyrics$title_word_count, lyrics$word_count)
```
 As expected, there is no correlation between these two quantitites.
 
 Let us fix the contracted words to their full forms first.

```{r}
# function to expand contractions in an English-language source
fix.contractions <- function(doc) {
  # "won't" is a special case as it does not expand to "wo not"
  doc <- gsub("won't", "will not", doc)
  doc <- gsub("can't", "can not", doc)
  doc <- gsub("n't", " not", doc)
  doc <- gsub("'ll", " will", doc)
  doc <- gsub("'re", " are", doc)
  doc <- gsub("'ve", " have", doc)
  doc <- gsub("'m", " am", doc)
  doc <- gsub("'d", " would", doc)
  # 's could be 'is' or could be possessive: it has no expansion
  doc <- gsub("'s", "", doc)
  return(doc)
}

# fix (expand) contractions
lyrics$text <- sapply(lyrics$text, fix.contractions)
```
 
 Remove special characters from lyrics
```{r}
# function to remove special characters
removeSpecialChars <- function(x) gsub("[^a-zA-Z0-9 ]", " ", x)
# remove special characters
lyrics$text <- sapply(lyrics$text, removeSpecialChars)
```

Convert all lyrics text to lower case
```{r}
# convert everything to lower case
lyrics$text <- sapply(lyrics$text, tolower)
```

Let's check the structure of one lyrics to see the changes.
```{r}
str(lyrics[13, ]$text, nchar.max = 300)
```
 
 SENTIMENT ANALYSIS OF LYRICS
 Let us perform sentiment analysis on the lyrics. There are various types of sentiment lexicons that can be used. Lets us have a look on them.
```{r}
library(tidytext)
library(tidyr)
get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
get_sentiments("loughran")
```

nrc seems to have large number of words and their sentiments compared to other two.

```{r}
nrc_sentiment <- get_sentiments("nrc")
unique(nrc_sentiment$sentiment)
```

Let us find the sentiments of each lyrics based on each NRC sentiments.

```{r}
lyrics_words <- select(lyrics, c("artist", "text"))
lyrics_words <- lyrics_words %>% unnest_tokens(word, text)
head(lyrics_words)
dim(lyrics_words)
```

Let's see words that depict "joy"
```{r}
joy <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "joy")) 
joy <- as.data.frame(sort(table(joy$word)))
columns_sentiment <- c("word", "Freq")
colnames(joy) <- columns_sentiment
tail(joy, 10)

```

Similarly, let's see other words that represents other 9 sentiments.

```{r}
trust <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "trust")) 
trust <- as.data.frame(sort(table(trust$word)))
colnames(trust) <- columns_sentiment
tail(trust, 10)
```

```{r}
fear <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "fear")) 
fear <- as.data.frame(sort(table(fear$word)))
colnames(fear) <- columns_sentiment

sadness <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "sadness")) 
sadness <- as.data.frame(sort(table(sadness$word)))
colnames(sadness) <- columns_sentiment

anger <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "anger")) 
anger <- as.data.frame(sort(table(anger$word)))
colnames(anger) <- columns_sentiment

surprise <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "surprise")) 
surprise <- as.data.frame(sort(table(surprise$word)))
colnames(surprise) <- columns_sentiment

disgust <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "disgust")) 
disgust <- as.data.frame(sort(table(disgust$word)))
colnames(disgust) <- columns_sentiment

anticipation <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "anticipation")) 
anticipation <- as.data.frame(sort(table(anticipation$word)))
colnames(anticipation) <- columns_sentiment
```

Let's plot the word occurences for each sentiment (except positive and negative)

```{r fig.width=8, fig.height=4, echo=FALSE}
p9 <- ggplot(data = tail(joy, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Joy") +
      guides(fill=FALSE) +
      tilt_theme
p10 <- ggplot(data = tail(trust, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Trust") +
      guides(fill=FALSE) +
      tilt_theme
p11 <- ggplot(data = tail(fear, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Fear") +
      guides(fill=FALSE) +
      tilt_theme
p12 <- ggplot(data = tail(sadness, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Sadness") +
      guides(fill=FALSE) +
      tilt_theme
p13 <- ggplot(data = tail(anger, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Anger") +
      guides(fill=FALSE) +
      tilt_theme
p14 <- ggplot(data = tail(surprise, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Surprise") +
      guides(fill=FALSE) +
      tilt_theme
p15 <- ggplot(data = tail(disgust, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Disgust") +
      guides(fill=FALSE) +
      tilt_theme
p16 <- ggplot(data = tail(anticipation, 10), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "white", size = 3) +
      ggtitle("Anticipation") +
      guides(fill=FALSE) +
      tilt_theme

multiplot(p9, p10, p11, p12, p13, p14, p15, p16, layout = matrix(c(1,2,3,4,5,6,7,8), nrow=2, byrow=TRUE))


```

Let's plot the wordcloud for each category

```{r fig.width=5, fig.height=8, echo=FALSE}

#Create two panels to add the word clouds to
par(mfrow=c(4,2))

wordcloud(words = joy$word, freq = joy$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Joy words", outer=FALSE)
          #colors= c("indianred1","indianred2","indianred3","indianred"))

wordcloud(words = trust$word, freq = trust$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Trust words", outer=FALSE)

wordcloud(words = fear$word, freq = fear$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Fear words", outer=FALSE)

wordcloud(words = sadness$word, freq = sadness$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Sadness words", outer=FALSE)

wordcloud(words = anger$word, freq = anger$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Anger words", outer=FALSE)

wordcloud(words = surprise$word, freq = surprise$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Surprise words", outer=FALSE)

wordcloud(words = disgust$word, freq = disgust$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Disgust words", outer=FALSE)

wordcloud(words = anticipation$word, freq = anticipation$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Anticipation words", outer=FALSE)
```

Let us check top words that depicts "positive" and "negative" sentiments in NRC Sentiment category.

```{r fig.width=5, fig.height=2, echo=FALSE}
pos <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "positive")) 
pos <- as.data.frame(sort(table(pos$word)))
colnames(pos) <- columns_sentiment

neg <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "negative")) 
neg <- as.data.frame(sort(table(neg$word)))
colnames(neg) <- columns_sentiment

p17 <- ggplot(data = tail(pos, 20), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3, angle=90) +
      ggtitle("Positive (NRC)") +
      guides(fill=FALSE) +
      tilt_theme
p18 <- ggplot(data = tail(neg, 20), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3, angle=90) +
      ggtitle("Negative (NRC)") +
      guides(fill=FALSE) +
      tilt_theme
multiplot(p17, p18, cols=2)
```




Let us check top words that depicts "positive" and "negative" sentiments in bing Sentiment category.

```{r fig.width=10, fig.height=5, echo=FALSE}
pos_b <- lyrics_words %>%
  inner_join(get_sentiments("bing") %>% 
  filter(sentiment == "positive")) 
pos_b <- as.data.frame(sort(table(pos_b$word)))
colnames(pos_b) <- columns_sentiment

neg_b <- lyrics_words %>%
  inner_join(get_sentiments("bing") %>% 
  filter(sentiment == "negative")) 
neg_b <- as.data.frame(sort(table(neg_b$word)))
colnames(neg_b) <- columns_sentiment

pos_l <- lyrics_words %>% 
  inner_join(get_sentiments("loughran") %>% 
  filter(sentiment == "positive")) 
pos_l <- as.data.frame(sort(table(pos_l$word)))
colnames(pos_l) <- columns_sentiment

neg_l <- lyrics_words %>%
  inner_join(get_sentiments("loughran") %>% 
  filter(sentiment == "negative")) 
neg_l <- as.data.frame(sort(table(neg_l$word)))
colnames(neg_l) <- columns_sentiment

p19 <- ggplot(data = tail(pos_b, 20), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3, angle=90) +
      ggtitle("Positive (BING)") +
      guides(fill=FALSE) +
      tilt_theme
p20 <- ggplot(data = tail(neg_b, 20), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3, angle=90) +
      ggtitle("Negative (BING)") +
      guides(fill=FALSE) +
      tilt_theme
p21 <- ggplot(data = tail(pos_l, 20), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3, angle=90) +
      ggtitle("Positive (LOUGHRAN)") +
      guides(fill=FALSE) +
      tilt_theme
p22 <- ggplot(data = tail(neg_l, 20), aes(word, Freq, fill = word)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3, angle=90) +
      ggtitle("Negative (LOUGHRAN)") +
      guides(fill=FALSE) +
      tilt_theme
multiplot(p19, p20,p21, p22,layout = matrix(c(1,2,3,4), nrow=2, byrow=TRUE))

```

It can be seen that positive and negative words are different for all three lexicons. 

Let's plot positive and negative word cloud


```{r fig.width=5, fig.height=7, echo=FALSE}

#Create two panels to add the word clouds to
par(mfrow=c(3,2))
#Create word cloud of positive words of NRC lexicon
wordcloud(words = pos$word, freq = pos$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Positive words in NRC Lexicon", outer=FALSE)
          #colors= c("indianred1","indianred2","indianred3","indianred"))
#Create word cloud of negative words of NRC lexicon
wordcloud(words = neg$word, freq = neg$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Accent"))
title("Negative words in NRC Lexicon", outer=FALSE)
#Create word cloud of positive words of bing lexicon
wordcloud(words = pos_b$word, freq = pos_b$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Positive words in Bing Lexicon", outer=FALSE)
#Create word cloud of negative words of bing lexicon
wordcloud(words = neg_b$word, freq = neg_b$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Accent"))
title("Negative words in Bing Lexicon", outer=FALSE)
#Create word cloud of positive words of loughran lexicon
wordcloud(words = pos_l$word, freq = pos_l$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
title("Positive words in Loughran Lexicon", outer=FALSE)
#Create word cloud of negative words of loughran lexicon
wordcloud(words = neg_l$word, freq = neg_l$Freq, min.freq = 50,
          max.words=100, scale=c(2, 1), random.order = FALSE, random.color = FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Accent"))
title("Negative words in Loughran Lexicon", outer=FALSE)
```


Let's check songs based on NRC sentiment (except positive and negative)

```{r}
lyrics_words <- select(lyrics, c("artist", "song","text"))
lyrics_words <- lyrics_words %>% unnest_tokens(word, text)
head(lyrics_words)
dim(lyrics_words)
```

```{r}
joy_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "joy")) 
joy_s <- as.data.frame(sort(table(joy_s$song)))
columns_sentiment <- c("Song", "Freq")
colnames(joy_s) <- columns_sentiment
tail(joy_s, 10)

trust_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "trust")) 
trust_s <- as.data.frame(sort(table(trust_s$song)))
colnames(trust_s) <- columns_sentiment
tail(trust_s, 10)

fear_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "fear")) 
fear_s <- as.data.frame(sort(table(fear_s$song)))
colnames(fear_s) <- columns_sentiment

sadness_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "sadness")) 
sadness_s <- as.data.frame(sort(table(sadness_s$song)))
colnames(sadness_s) <- columns_sentiment

anger_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "anger")) 
anger_s <- as.data.frame(sort(table(anger_s$song)))
colnames(anger_s) <- columns_sentiment

surprise_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "surprise")) 
surprise_s <- as.data.frame(sort(table(surprise_s$song)))
colnames(surprise_s) <- columns_sentiment

disgust_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "disgust")) 
disgust_s <- as.data.frame(sort(table(disgust_s$song)))
colnames(disgust_s) <- columns_sentiment

anticipation_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "anticipation")) 
anticipation_s <- as.data.frame(sort(table(anticipation_s$song)))
colnames(anticipation_s) <- columns_sentiment
```

Let's plot the top 10 songs in each category


```{r fig.width=8, fig.height=4, echo=FALSE}
p23 <- ggplot(data = tail(joy_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Joy") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p24 <- ggplot(data = tail(trust_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Trust") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p25 <- ggplot(data = tail(fear_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Fear") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p26 <- ggplot(data = tail(sadness_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Sadness") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p27 <- ggplot(data = tail(anger_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Anger") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p28 <- ggplot(data = tail(surprise_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Surprise") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p29 <- ggplot(data = tail(disgust_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Disgust") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p30 <- ggplot(data = tail(anticipation_s, 10), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), vjust = 1.6, color = "black", size = 3) +
      ggtitle("Anticipation") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme

multiplot(p23, p24, p25, p26, p27, p28, p29,p30, layout = matrix(c(1,2,3,4,5,6,7,8), nrow=2, byrow=TRUE))

```



Let's do the same for Positive and Negative sentiment in NRC lexicon cateogory.
```{r fig.width=5, fig.height=2, echo=FALSE}
pos_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "positive")) 
pos_s <- as.data.frame(sort(table(pos_s$song)))
colnames(pos_s) <- columns_sentiment

neg_s <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "negative")) 
neg_s <- as.data.frame(sort(table(neg_s$song)))
colnames(neg_s) <- columns_sentiment

p31 <- ggplot(data = tail(pos_s, 20), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Positive (NRC)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p32 <- ggplot(data = tail(neg_s, 20), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq) , color = "black", size = 3) +
      ggtitle("Negative (NRC)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
multiplot(p31, p32, cols=2)
```

Let us find positive and negative songs based on bing and loughran lexicon cateogories.

```{r fig.width=10, fig.height=5, echo=FALSE}
pos_bs <- lyrics_words %>%
  inner_join(get_sentiments("bing") %>% 
  filter(sentiment == "positive")) 
pos_bs <- as.data.frame(sort(table(pos_bs$song)))
colnames(pos_bs) <- columns_sentiment

neg_bs <- lyrics_words %>%
  inner_join(get_sentiments("bing") %>% 
  filter(sentiment == "negative")) 
neg_bs <- as.data.frame(sort(table(neg_bs$song)))
colnames(neg_bs) <- columns_sentiment

pos_ls <- lyrics_words %>% 
  inner_join(get_sentiments("loughran") %>% 
  filter(sentiment == "positive")) 
pos_ls <- as.data.frame(sort(table(pos_ls$song)))
colnames(pos_ls) <- columns_sentiment

neg_ls <- lyrics_words %>%
  inner_join(get_sentiments("loughran") %>% 
  filter(sentiment == "negative")) 
neg_ls <- as.data.frame(sort(table(neg_ls$song)))
colnames(neg_ls) <- columns_sentiment

p33 <- ggplot(data = tail(pos_bs, 20), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Positive (BING)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p34 <- ggplot(data = tail(neg_bs, 20), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Negative (BING)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p35 <- ggplot(data = tail(pos_ls, 20), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Positive (LOUGHRAN)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p36 <- ggplot(data = tail(neg_ls, 20), aes(Song, Freq, fill = Song)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Negative (LOUGHRAN)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
multiplot(p33, p34,p35, p36,layout = matrix(c(1,2,3,4), nrow=2, byrow=TRUE))

```

The positive and negative songs in each category are different. This is something known already, since the positive words and negative words in ech categories were also different.

The same analysis can be done for artists. Let's check out artists who write most songs in each NRC lexicon sentiment cateogry.

```{r}
joy_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "joy")) 
joy_a <- as.data.frame(sort(table(joy_a$artist)))
columns_sentiment <- c("artist", "Freq")
colnames(joy_a) <- columns_sentiment
tail(joy_a, 10)

trust_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "trust")) 
trust_a <- as.data.frame(sort(table(trust_a$artist)))
colnames(trust_a) <- columns_sentiment
tail(trust_a, 10)

fear_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "fear")) 
fear_a <- as.data.frame(sort(table(fear_a$artist)))
colnames(fear_a) <- columns_sentiment

sadness_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "sadness")) 
sadness_a <- as.data.frame(sort(table(sadness_a$artist)))
colnames(sadness_a) <- columns_sentiment

anger_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "anger")) 
anger_a <- as.data.frame(sort(table(anger_a$artist)))
colnames(anger_a) <- columns_sentiment

surprise_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "surprise")) 
surprise_a <- as.data.frame(sort(table(surprise_a$artist)))
colnames(surprise_a) <- columns_sentiment

disgust_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "disgust")) 
disgust_a <- as.data.frame(sort(table(disgust_a$artist)))
colnames(disgust_a) <- columns_sentiment

anticipation_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "anticipation")) 
anticipation_a <- as.data.frame(sort(table(anticipation_a$artist)))
colnames(anticipation_a) <- columns_sentiment
```

Let's plot the top 10 artists in each category


```{r fig.width=8, fig.height=4, echo=FALSE}
p37 <- ggplot(data = tail(joy_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Joy") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p38 <- ggplot(data = tail(trust_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Trust") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p39 <- ggplot(data = tail(fear_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Fear") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p40 <- ggplot(data = tail(sadness_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Sadness") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p41 <- ggplot(data = tail(anger_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Anger") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p42 <- ggplot(data = tail(surprise_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Surprise") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p43 <- ggplot(data = tail(disgust_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq),color = "black", size = 3) +
      ggtitle("Disgust") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p44 <- ggplot(data = tail(anticipation_a, 10), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Anticipation") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme

multiplot(p37, p38, p39, p40, p41, p42, p43, p44, layout = matrix(c(1,2,3,4,5,6,7,8), nrow=2, byrow=TRUE))

```

Let's do the same for Positive and Negative sentiment in NRC lexicon cateogory.
```{r fig.width=5, fig.height=2, echo=FALSE}
pos_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "positive")) 
pos_a <- as.data.frame(sort(table(pos_a$artist)))
colnames(pos_a) <- columns_sentiment

neg_a <- lyrics_words %>%
  inner_join(get_sentiments("nrc") %>% 
  filter(sentiment == "negative")) 
neg_a <- as.data.frame(sort(table(neg_a$artist)))
colnames(neg_a) <- columns_sentiment

p45 <- ggplot(data = tail(pos_a, 15), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Positive (NRC)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p46 <- ggplot(data = tail(neg_a, 15), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq) , color = "black", size = 3) +
      ggtitle("Negative (NRC)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
multiplot(p45, p46, cols=2)
```

Let us find positive and negative songs based on bing and loughran lexicon cateogories.

```{r fig.width=10, fig.height=5, echo=FALSE}
pos_ba <- lyrics_words %>%
  inner_join(get_sentiments("bing") %>% 
  filter(sentiment == "positive")) 
pos_ba <- as.data.frame(sort(table(pos_ba$artist)))
colnames(pos_ba) <- columns_sentiment

neg_ba <- lyrics_words %>%
  inner_join(get_sentiments("bing") %>% 
  filter(sentiment == "negative")) 
neg_ba <- as.data.frame(sort(table(neg_ba$artist)))
colnames(neg_ba) <- columns_sentiment

pos_la <- lyrics_words %>% 
  inner_join(get_sentiments("loughran") %>% 
  filter(sentiment == "positive")) 
pos_la <- as.data.frame(sort(table(pos_la$artist)))
colnames(pos_la) <- columns_sentiment

neg_la <- lyrics_words %>%
  inner_join(get_sentiments("loughran") %>% 
  filter(sentiment == "negative")) 
neg_la <- as.data.frame(sort(table(neg_la$artist)))
colnames(neg_la) <- columns_sentiment

p47 <- ggplot(data = tail(pos_ba, 15), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Positive (BING)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p48 <- ggplot(data = tail(neg_ba, 15), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Negative (BING)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p49 <- ggplot(data = tail(pos_la, 15), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Positive (LOUGHRAN)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p50 <- ggplot(data = tail(neg_la, 15), aes(artist, Freq, fill = artist)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Negative (LOUGHRAN)") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
multiplot(p47, p48,p49, p50,layout = matrix(c(1,2,3,4), nrow=2, byrow=TRUE))

```

Let's check out bigrams in the songs lyrics

```{r echo=FALSE}
lyrics_bigram <- unnest_tokens(lyrics, input = text, output = bigram, token = "ngrams", n=2)


bigram_filtered<-lyrics_bigram %>%separate(bigram,c("word1","word2",sep=" "))%>%
filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
  
 
  bigram_united <- bigram_filtered %>%
  unite(bigram, word1, word2, sep = " ")
head(bigram_united)


```


```{r}
bigram_counts <- as.data.frame(sort(table(bigram_united$bigram)))
colnames(bigram_counts) <- c("bigrams", "Freq")
bigram_counts_head <- bigram_counts %>% arrange(desc(Freq))%>% head(15)
bigram_counts_head

p51 <- ggplot(data = bigram_counts_head, aes(bigrams, Freq, fill = bigrams)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Top 15 bigrams in songs") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p51
```

```{r}
 wordcloud(bigram_counts$bigrams,bigram_counts$Freq,max.words=100,random.order = F,colors=brewer.pal(8, "Dark2"))
title(main="Bigram wordcloud")
```

Let's check out trigrams in the lyrics of songs.
```{r echo=FALSE}
lyrics_trigram <- unnest_tokens(lyrics, input = text, output = trigram, token = "ngrams", n=3)


trigram_filtered<-lyrics_trigram %>%separate(trigram,c("word1","word2","word3",sep=" "))%>%
filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  filter(!word3 %in% stop_words$word)
  
 
  trigram_united <- trigram_filtered %>%
  unite(trigram, word1, word2,word3, sep = " ")
```  

```{r}
trigram_counts <- as.data.frame(sort(table(trigram_united$trigram)))
colnames(trigram_counts) <- c("trigrams", "Freq")
trigram_counts_head <- trigram_counts %>% arrange(desc(Freq))%>% head(15)
trigram_counts_head

p52 <- ggplot(data = trigram_counts_head, aes(trigrams, Freq, fill = trigrams)) +
      geom_bar(position = "dodge", stat = "identity") +
      geom_text(aes(label = Freq), color = "black", size = 3) +
      ggtitle("Top 15 trigrams in songs") +
      guides(fill=FALSE) + coord_flip() +
      tilt_theme
p52
 
wordcloud(trigram_counts$trigrams,trigram_counts$Freq,max.words=100,random.order = F,colors=brewer.pal(8, "Dark2"))
title(main="Trigram wordcloud")
```

Let us find out top words used by any artist in their songs.

```{r}
get_wordcloud_of_artist_songs <- function(artist_name){
artist_songs <- filter(lyrics, lyrics$artist == artist_name)
texts <- artist_songs$text
corpus <- Corpus(VectorSource(texts))
corpus <- tm_map(corpus, PlainTextDocument)
corpus <- Corpus(VectorSource(corpus))
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,scale=c(2,1),
          max.words=100, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
}
```

```{r}
get_wordcloud_of_artist_songs("ABBA")
```

```{r}
get_wordcloud_of_artist_songs("Rihanna")
```

```{r}
get_wordcloud_of_artist_songs("Usher")
```


